Supplementary Information

### 2015-08-27 manuscript/AC-reply-to-referees.tex

  \begin{refquote}

    My concern here is how much of the pattern can be attributed to true
    ecological effects – as opposed to the consequences of differences in the
    effort used to describe different food webs... It could be explored to some extent by
    analysing for an effect of particular authors (e.g. the Thompson and
    Townsend food webs which dominate the stream dataset) but ultimately there
    is every chance that the results here are artefacts of differences in
    sampling effort between studies. Might it be possible to randomise the
    datasets, excluding particular food webs (or authors) to determine
    robustness to this effect??

  \end{refquote}

  \textbf{R:} We share the reviewer's concern that variation in sampling
  effort could be driving the trends we observe, as it is highly likely that
  the authors of all 196 webs in our dataset did not follow the same procedures.
  Following the reviewer's suggestion, we have jackknifed our models by systematically
  removing each web in turn and then re-running our analysis. We then repeated this
  process, removing each author (or set of authors where authors worked on identical sets
  of food webs) in turn and re-fitting the models. We are happy to report that removing
  any single web had no significant effect on any model, nor did removing any author
  (we removed Thompson and Townsend separately as each worked on one web that the other
  did not). We have added a description of these methods and summary figures of their results
  to the supporting information, and we thank the reviewer for suggesting this validation
  of our overall results. We have also added a brief note in the main text to direct other
  readers with similar concerns to the supporting information (lines XX-XX). [[At the end of the 
  section describing our analyses - didn't seeem to fit after the first mention of sampling effort]]
  [[done]]

@@

  \begin{quotation}

    As a supplemental check to ensure that variation in sampling effor across
    food webs was not responsible for the trends we observed, we then repeated
    our analyses using jackknifed data sets in which we 1) sequentially removed
    each food web in the dataset and 2) sequentially removed sets of food webs 
    associated with each author of the food webs in the dataset. Parameter 
    estimates varied very little across either series of jackknifes (\emph{Appendix SXX}).

@@

  \end{quotation}

  \newpage

% -----------------------------------------------------------------------------
% -----------------------------------------------------------------------------
{\Large \bf Reply to Reviewer 2}
% ---------------------------------------

  The Reviewer stated that our study is interesting and covers ``a timely and
  important topic. The core analysis of the paper (scaling relationships of
  food web metrics with species richness) are convincingly presented and
  comprehensively discussed.'' Nevertheless, the Reviewer has suggested several
  ways in which our manuscript could be improved. We have responded to each suggestion
  in turn, and believe that our replies (preceded by \textbf{R}) have further improved
  the manuscript.

@@

    \smallskip

     For readers outside the field, please explain in more detail the
     difference between `original species' and `trophic species' in food webs.

  \end{refquote}

  \textbf{R:} We thank the reviewer for their concern that our manuscript
  should be accessible to as wide a variety of researchers as possible, and
  have followed the reviewer's suggestions in several places in the manuscript.
  We now include a conceptual figure to clearly and succinctly link the changes
  to scaling relationships we expected to the two versions of the latitude-niche
  breadth hypothesis (Fig. 1).

  At lines XX-XX we have added parenthetical explanations of the original and 
  trophic-species versions of food webs. We have also added parenthetical definitions
  of link density, generality, and vulnerability at the begining of the results section
  to remind readers of the biological meaning of these measures.

@@

  \begin{quotation}

    We therefore analysed both original (i.e., without aggregating any species) and
    trophic-species (i.e., after aggregating species with identical predators and
    prey) versions of the dataset ...

      Link density (mean number of feeding links per species), generality
      (mean number of prey per species), and vulnerability (mean number of predators
      per species) were strongly and positively correlated ...

@@

     \smallskip

      The start of the Discussion confused me since it presents another
      additional analysis. This analysis on the latitudinal gradient in
      species richness is interesting and important for the interpretation of
      the patterns, but should already be presented in Methods and Results
      section.

  \end{refquote}

  \textbf{R:} We had originally viewed this analysis as a very minor part of our paper,
  but are happy to follow the Reviewer's suggestion and integrate it in the methods and
  results along with our main analyses. We have described the methods for this analysis
  in lines XX-XX and summarize our results in lines XX-XX.

    \subsection*{Gradients over latitude}

    To put our dataset in the context of other research on latitudinal gradients in species richness,
    we first examined simple linear relationships between latitude and each of 
    species richness, links per species, generality, vulnerability, and proportions
    of basal resources, intermediate consumers, and top predators. We fit models of the form

    \begin{equation}
    \label{Latfull}
    S_{i} = \alpha_{0} + \alpha_{1} L_{i} + \alpha_{2} E_{i} + \alpha_{3} L_{i} E_{i} + \epsilon_{i} ,
    \end{equation}

    \noindent where $S_{i}$ is the species richness of web $i$, $L_{i}$ its absolute
    latitude (degrees north or south  regardless of direction), $E_{i}$ is a categorical
    variable indicating the ecosystem type of network $i$ (comprising terms for stream, 
    marine, lake, and terrestrial networks with estuarine
    networks providing the intercept) and $\epsilon_{i}$ is a residual error term. 
    We next calculated the AIC
    of the maximal model as well as the AIC's of a suite of candidate simplified models identified
    using the R~\citep{R} function dredge from package MuMIn~\citep{MuMIn}. 
    Simplified models were obtained by
    systematically removing all possible combinations of terms from the full model.
    The best-fitting model was then determined to be the model with the fewest terms 
    where $\Delta$AIC\textless2. If several models shared the fewest number of terms 
    and had $\Delta$AIC\textless2, the model with the lowest AIC in that set was chosen as the best-fit
    model.

    Contrary to the expected latitudinal gradient, the best-fit version of
    equation~(\ref{Latfull}) for species richness did not  include a signficant
    effect of latitude for any  ecosystem type. Further, there were no
    relationships between link density, generality, or vulnerability with
    latitude for any ecosystem type.

@@

  \begin{refquote}

    Please describe in more detail why you used the chosen habitat categories
    and provide the number of webs per habitat type also in the main
    manuscript.

  \end{refquote}

  \textbf{R:} We followed the habitat categories given by the publications which
  aggregated the food webs. Although more detailed descriptions were also provided, 
  we opted to use the set of five more general habitat types in order to have a 
  greater number of food webs in each category.

  We have added a brief explanation, and the number of webs in each ecosystem type, at lines XX-XX:

  \begin{quotation}
    We grouped food web by ecosystem type (stream, N=71; lake, N=47; marine, N=28; estuarine, N=18;
    and terrestrial, N=31) according to their designation in previous aggregations of food webs 
    (i.e.,~\citet{GlobalWeb,Riede2011,Dunne2013}).

### 2015-08-27 manuscript/latitudepaper.tex
Significant additions to the reply

    We compiled a list of 166 empirical food webs from
    multiple sources (see \emph{Appendix S1} for web origins and selection
    criteria). We recorded study site latitude from the original source where
    possible or, where study sites were described but exact coordinates were not
    given, obtained estimated coordinates using Google Earth~\citep{GoogleEarth}.
    We then divided the species in each web into basal resources (those species with
    consumers but no prey), top predators (those species with prey but no predators),
    and intermediate consumers (species with both predators and prey, including cannibalistic
    species).

    As the food webs in this dataset are derived from a variety of sources and were compiled over many decades, it
    is likely that they vary in their resolution and in the amount of sampling effort invested in their assembly.
    Many analyses of food-web structure attempt to reduce this variation by aggregating species with identical predator and prey
    sets to form ``trophic species'' webs ~\citep{Martinez1991,Dunne2004,Vermaat2009,Dunne2013}. As our study
    is concerned directly with the number of species at a particular latitude, however, we did not wish to ignore 
    species with identical sets of interactions. We therefore analysed both original and trophic-species versions
    of the dataset; in each case using the number of species and 
    links in each web to calculate the mean link density (number of links per species), mean generality 
    (number of prey per species), and mean vulnerability (number of predators per species) of the web. 
    The version of the dataset used did not qualitatively change the results, suggesting that
    the scaling relationships between species richness, other food-web properties, and latitude are very 
    similar whether or not redundant species are included. For simplicity, here we present only
    the results for the original webs.

@@

    We then simplified each model. To do this, we calculated the AIC of the
    maximal model as well as the AIC's of a suite of candidate simplified
    models identified using the R~\citep{R} function dredge from package
    MuMIn~\citep{MuMIn}.  Simplified models were obtained by systematically
    removing all possible combinations  of terms from the full model with the
    restriction that species richness was retained in all reduced models. The
    best-fitting model was then determined to be the model with the fewest
    terms  where $\Delta$AIC\textless2, as this model is the least likely to suffer from over-fitting.
    % If several models shared the fewest
    % number of terms  and had $\Delta$AIC\textless2, the model with the lowest
    % AIC in that set was chosen as the best-fit model.

@@

  Link density, generality, and vulnerability were strongly and positively
  correlated ($R^2$=0.891 for link density and generality,
  $R^2$\textgreater0.999 for link density and vulnerability, and $R^2$=0.890
  for generality and vulnerability). Each property 
  increased with increasing
  species richness ($\beta_0$=0.637, $p$\textless0.001; $\beta_0$=0.553,
  $p$\textless0.001; and $\beta_0$=0.637, $p$\textless0.001, respectively;
  Fig~\ref{props_v_lat}).For estuarine, marine, and terrestrial food webs the
  strength of this scaling did not vary with latitude
  ($\beta_{Latitude}$=-0.001, $p$=0.365 for link density;
  $\beta_{Latitude}$=-0.001, $p$=0.535 for generality; and
  $\beta_{Latitude}$=-0.001, $p$=0.366 for vulnerability; Fig.~\ref{S}). In
  lake food webs, however, the scaling of each property was stronger towards
  the poles ($\beta_{Latitude:Lake}$=0.004, $p$=0.019;
  $\beta_{Latitude:Lake}$=0.005, $p$=0.004; and
  $\beta_{Latitude:Lake}$=0.004, $p$=0.018, respectively). In stream food
  webs, generality increased more rapidly towards the poles
  ($\beta_{Latitude:Stream}$=0.007, $p$=0.001) while link density and
  vulnerability did not vary with latitude (i.e., the interaction term $\beta_{Latitude:Stream}$ was not retained in the best-fit models).

>>>>

    We compiled a list of 166 empirical food webs from
    multiple sources (see \emph{Appendix S1} for web origins and selection
    criteria). We recorded study site latitude from the original source where
    possible or, where study sites were described but exact coordinates were not
    given, obtained estimated coordinates using Google Earth~\citep{GoogleEarth}.
    We then divided the species in each web into basal resources (those species with
    consumers but no prey), top predators (those species with prey but no predators),
    and intermediate consumers (species with both predators and prey, including cannibalistic
    species). We grouped food web by ecosystem type (stream, N=71; lake, N=47; marine, 
    N=28; estuarine, N=18; and terrestrial, N=31) according to their designation in 
    previous aggregations of food webs (i.e.,~\citet{GlobalWeb,Riede2011,Dunne2013}).

    As the food webs in this dataset are derived from a variety of sources and were compiled over many decades, it
    is likely that they vary in their resolution and in the amount of sampling effort invested in their assembly.
    Many analyses of food-web structure attempt to reduce this variation by aggregating species with identical predator and prey
    sets to form ``trophic species'' webs ~\citep{Martinez1991,Dunne2004,Vermaat2009,Dunne2013}. As our study
    is concerned directly with the number of species at a particular latitude, however, we did not wish to ignore 
    species with identical sets of interactions. We therefore analysed both original (i.e., without aggregating 
    any species) and trophic-species (i.e., after aggregating species with identical predators and prey) versions
    of the dataset; in each case using the number of species and 
    feeding links in each web to calculate the mean link density (number of links per species), mean generality 
    (number of prey per species), and mean vulnerability (number of predators per species) of the web. 
    The version of the dataset used did not qualitatively change the results, suggesting that
    the scaling relationships between species richness, other food-web properties, and latitude are very 
    similar whether or not redundant species are included. For simplicity, here we present only
    the results for the original (un-aggregated) webs.

  \subsection*{Gradients over latitude}

    To put our dataset in the context of other research on latitudinal gradients in species richness,
    we first examined simple linear relationships between latitude and each of 
    species richness, links per species, generality, vulnerability, and proportions
    of basal resources, intermediate consumers, and top predators. We fit models of the form

    \begin{equation}
    \label{Latfull}
    S_{i} = \alpha_{0} + \alpha_{1} L_{i} + \alpha_{2} E_{i} + \alpha_{3} L_{i} E_{i} + \epsilon_{i} ,
    \end{equation}

    \noindent where $S_{i}$ is the species richness of web $i$, $L_{i}$ its absolute
    latitude (degrees north or south  regardless of direction), $E_{i}$ is a categorical
    variable indicating the ecosystem type of network $i$ (comprising terms for stream, 
    marine, lake, and terrestrial networks with estuarine
    networks providing the intercept) and $\epsilon_{i}$ is a residual error term. 
    We next calculated the AIC
    of the maximal model as well as the AIC's of a suite of candidate simplified models identified
    using the R~\citep{R} function dredge from package MuMIn~\citep{MuMIn}. 
    Simplified models were obtained by
    systematically removing all possible combinations of terms from the full model.
    The best-fitting model was then determined to be the model with the fewest terms 
    where $\Delta$AIC\textless2. If several models shared the fewest number of terms 
    and had $\Delta$AIC\textless2, the model with the lowest AIC in that set was chosen as the best-fit
    model.

@@

    We then simplified each model. To do this, we calculated the AIC of the
    maximal model as well as the AIC's of a suite of candidate simplified
    models identified using the R~\citep{R} function dredge from package
    MuMIn~\citep{MuMIn}.  Simplified models were obtained by systematically
    removing all possible combinations  of terms from the full model with the
    restriction that species richness was retained in all reduced models. The
    best-fitting model was then determined to be the model with the fewest
    terms  where $\Delta$AIC\textless2, as this model is the least likely to suffer from over-fitting.
    As a supplemental check to ensure that variation in sampling effor across
    food webs was not responsible for the trends we observed, we then repeated
    our analyses using jackknifed data sets in which we 1) sequentially removed
    each food web in the dataset and 2) sequentially removed sets of food webs 
    associated with each author of the food webs in the dataset. Parameter 
    estimates varied very little across either series of jackknifes (\emph{Appendix SXX}).
    % If several models shared the fewest
    % number of terms  and had $\Delta$AIC\textless2, the model with the lowest
    % AIC in that set was chosen as the best-fit model.

@@

  Contrary to the expected latitudinal gradient, the best-fit version of
  equation~(\ref{Latfull}) for species richness did not  include a signficant
  effect of latitude for any  ecosystem type. Further, there were no
  relationships between link density, generality, or vulnerability with
  latitude for any ecosystem type. Link density (mean number of feeding links
  per species), generality (mean number of prey per species), and
  vulnerability (mean number of predators per species) were strongly and
  positively correlated ($R^2$=0.891 for link density and generality,
  $R^2$\textgreater0.999 for link density and vulnerability, and $R^2$=0.890
  for generality and vulnerability).

  Each property increased with increasing
  species richness ($\beta_0$=0.637, $p$\textless0.001; $\beta_0$=0.553,
  $p$\textless0.001; and $\beta_0$=0.637, $p$\textless0.001, respectively;
  Fig~\ref{props_v_lat}).For estuarine, marine, and terrestrial food webs the
  strength of this scaling did not vary with latitude
  ($\beta_{Latitude}$=-0.001, $p$=0.365 for link density;
  $\beta_{Latitude}$=-0.001, $p$=0.535 for generality; and
  $\beta_{Latitude}$=-0.001, $p$=0.366 for vulnerability; Fig.~\ref{S}). In
  lake food webs, however, the scaling of each property was stronger towards
  the poles ($\beta_{Latitude:Lake}$=0.004, $p$=0.019;
  $\beta_{Latitude:Lake}$=0.005, $p$=0.004; and
  $\beta_{Latitude:Lake}$=0.004, $p$=0.018, respectively). In stream food
  webs, generality increased more rapidly towards the poles
  ($\beta_{Latitude:Stream}$=0.007, $p$=0.001) while link density and
  vulnerability did not vary with latitude (i.e., the interaction term $\beta_{Latitude:Stream}$ was not retained in the best-fit models).

### 2015-08-27 manuscript/supplement.tex
Significant additions to the reply

\subsection*{Appendix S3: Relationships with latitude}

    To determine whether there were latitudinal gradients in food-web structure,
    we first examined simple linear relationships between latitude and each of 
    species richness, links per species, generality, vulnerability, and proportions
    of basal resources, intermediate consumers, and top predators. We fit models of the form

    \begin{equation}
    \label{Latfull}
    S_{i} = \alpha_{0} + \alpha_{1} L_{i} + \alpha_{2} E_{i} + \alpha_{3} L_{i} E_{i} + \epsilon_{i} ,
    \end{equation}

    \noindent where $S_{i}$ is the species richness of web $i$, $L_{i}$ its absolute
    latitude (degrees north or south  regardless of direction), $E_{i}$ is a categorical
    variable indicating the ecosystem type of network $i$ (comprising terms for stream, 
    marine, lake, and terrestrial networks with estuarine
    networks providing the intercept) and $\epsilon_{i}$ is a residual error term. 
    We next calculated the AIC
    of the maximal model as well as the AIC's of a suite of candidate simplified models identified
    using the R~\citep{R} function dredge from package MuMIn~\citep{MuMIn}. 
    Simplified models were obtained by
    systematically removing all possible combinations of terms from the full model.
    The best-fitting model was then determined to be the model with the fewest terms 
    where $\Delta$AIC\textless2. If several models shared the fewest number of terms 
    and had $\Delta$AIC\textless2, the model with the lowest AIC in that set was chosen as the best-fit
    model.

  \subsubsection*{Results}
    Contrary to the expected latitudinal gradient, the best-fit version of
    equation~(\ref{Latfull}) for species richness did not 
    include a signficant effect of latitude for any 
    ecosystem type. Further, there were no relationships
    between link density, generality, or vulnerability
    with latitude for any ecosystem type.

    The relationships between latitude and the proportions of species at each trophic level were more complex. 
    The proportion of basal resources increased towards the poles in stream webs ($\beta_{Latitude:stream}$=0.011, $p$\textless0.001) 
    but did not vary with latitude in estuarine, marine, lake, or terrestrial 
    food webs ($\beta_{Latitude}$=0.001, $p$=0.080),
    while the proportion of intermediate consumers decreased towards the poles in all ecosystem types
    except lakes ($\beta_{Latitude}$=-0.003, $p$=0.015 and
    $\beta_{Latitude:Lake}$=0.005, $p$=0.016).
    The proportion of top predators in a food web, 
    meanwhile, decreased towards the poles in lake and 
    stream food webs but did not vary in other ecosystem
    types ($\beta_{Latitude}$=0.002, $p$=0.151; 
    $\beta_{Latitude:Lake}$=-0.006, $p$=0.001; and
    $\beta_{Latitude:stream}$=-0.007, $p$=0.006).

    % Does this need to be taken into account? Do I mention
    % it later?
    % All of these trends were robust to the removal of outliers.
    % Was it? Do I even care?

@@

>>>>

@@

    The relationships between latitude and the proportions of species at each
    trophic level were more complex than the flat relationship between latitude
    and species richness in this dataset.  The proportion of basal resources
    increased towards the poles in stream webs
    ($\beta_{Latitude:stream}$=0.011, $p$\textless0.001)  but did not vary
    with latitude in estuarine, marine, lake, or terrestrial  food webs
    ($\beta_{Latitude}$=0.001, $p$=0.080), while the proportion of
    intermediate consumers decreased towards the poles in all ecosystem types
    except lakes ($\beta_{Latitude}$=-0.003, $p$=0.015 and
    $\beta_{Latitude:Lake}$=0.005, $p$=0.016). The proportion of top predators
    in a food web,  meanwhile, decreased towards the poles in lake and  stream
    food webs but did not vary in other ecosystem types
    ($\beta_{Latitude}$=0.002, $p$=0.151;  $\beta_{Latitude:Lake}$=-0.006,
    $p$=0.001; and $\beta_{Latitude:stream}$=-0.007, $p$=0.006).
